### 2018-04-11 submission/redraft.rmd
Add paragraph

---
title: "Automated detection of data fabrication using statistical tools"
author: "Chris HJ Hartgerink, Jan G Voelkel, Jelte M Wicherts, Marcel ALM van Assen"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
 pdf_document:
  toc: yes
 html_document:
  toc: yes
  toc_depth: 2
csl: ../bibliography/apa.csl
bibliography: ../bibliography/library2.bib
---

```{r echo = FALSE}
source('../functions/digit_analysis.R')
source('../functions/std_var.R')
source('../functions/fisher_method.R')
```

# Introduction

```{r echo = FALSE}
m <- 50
sd <- 10
n <- 100

set.seed(1234)
x1_1 <- replicate(n = 10, expr = rnorm(n, m, sd))
x1_2 <- replicate(n = 10, expr = rnorm(n, m, sd))

ex1_1em <- apply(x1_1, 2, mean)
ex1_1esd <- apply(x1_1, 2, sd)
ex1_1cm <- apply(x1_2, 2, mean)
ex1_1csd <- apply(x1_2, 2, sd)
ex1_1tp <- NULL
for (i in 1:dim(x1_1)[2]) ex1_1tp[i] <- t.test(x = x1_1[,i], y = x1_2[,i], var.equal=TRUE)$p.value
ex1_1tp <- unlist(ex1_1tp)

set.seed(1234)
x2_1em <- replicate(n = 10, expr = m + runif(1, 0, 20))
x2_2cm <- replicate(n = 10, expr = m + runif(1, 0, 20))
x2_1esd <- replicate(n = 10, expr = sd + runif(1, 0, 1.5))
x2_2csd <- replicate(n = 10, expr = sd + runif(1, 0, 1.5))

sdpooled <- ((n - 1) * x2_1esd^2 + (n - 1) * x2_2csd^2)/(n * 2 - 2)
tval <- (x2_1em - x2_2cm) / sdpooled
ex2_1tp <- pt(abs(tval), n * 2 - 2, lower.tail = FALSE) * 2

# Just saving for later use in theory section
# set.seed(1234)
# std_var(n = rep(100, length(x2_1esd) * 2), sds = c(x2_1esd, x2_2csd), iter = 10000)

df <- data.frame(study = sprintf("Study %s", 1:10),
         sprintf("%s (%s)", round(ex1_1em, 3), round(ex1_1esd, 3)),
         sprintf("%s (%s)", round(ex1_1cm, 3), round(ex1_1csd, 3)),
         ex1_1tp,
         sprintf("%s (%s)", round(x2_1em, 3), round(x2_1esd, 3)),
         sprintf("%s (%s)", round(x2_2cm, 3), round(x2_2csd, 3)),
         ex2_1tp)
names(df) <- c("Study",
        "$M_E$ ($SD_E$) [S1]",
        "$M_C$ ($SD_C$) [S1]",
        "P-value [S1]",
        "$M_E$ ($SD_E$) [S2]",
        "$M_C$ ($SD_C$) [S2]",
        "P-value [S2]")
```

<!-- consistency of data fabrication v data anomalies -->

Any field of empirical inquiry is faced with cases of scientific misconduct at some point, either in the form of fabrication, falsification or plagiarism (FFP). 
Psychology faced Stapel; medical sciences faced Poldermans and Macchiarini; life sciences faced Voignet; physical sciences faced Sch&ouml;n --- these are just a few examples of misconduct cases in the last decade. 
Overall, an estimated 2\% of all scholars admit to falsifying or fabricating research results at least once [@10.1371/journal.pone.0005738], which due to its self-report nature is likely to be an underestimate. 
The detection rate of data fabrication is likely to be even lower; for example, only around a dozen cases become public in the United States and the Netherlands, despite that there are several hundreds of thousands of researchers in these countries. 
At best, this suggests a detection rate below 1\% of those 2\% who admit to fabricating data --- the tip of a seemingly much larger iceberg.

In order to stifle attempts at data fabrication, improved detection of fabricated data is considered to deter such behavior. 
This idea is based on deterrence theory [@leviathan], which stipulates that increased risk of detection decreases the utility of scientific misconduct, hence, fewer people will engage in it. 
Detection techniques have developed to varying degrees for fabrication, falsification, and plagiarism.
Plagiarism scanners have been around the longest [e.g., @10.1109/13.28038] and are widely implemented not only at journals but also in student evaluation. 
For data fabrication, developments around detecting image manipulation are more recent and these methods are being implemented at journals. 
For example, the Journal of Cell Biology and the EMBO journal scan each submitted image for potential manipulation [@The_Journal_of_Cell_Biology2015-vh;@10.1038/546575a], which supposedly increases the risk of (blatant) image manipulation. 
More recently, algorithms are being developed to automate the scanning of images for such manipulations [@10.1007/s11948-016-9841-7]. 
Moreover, the application of such tools can also help researchers systematically evaluate research articles in order to estimate the extent to which image manipulation occurs [4\% of all papers are estimated to contain manipulated images, @10.1128/mBio.00809-16] or what factors are predictive of image manipulation [@10.1007/s11948-018-0023-7].

Detection methods for data fabrication in empirical research are often based on a mix of psychology theory and statistics theory. 
Because humans are notoriously bad at understanding and estimating randomness [@10.1126/science.185.4157.1124;@10.1037/h0031322;@10.1037/1082-989X.5.2.241], this could manifest itself in the fundamentally probabilistic data they try to fabricate. 
As such, when data are fabricated, principles of statistics and randomness could easily be violated at various dimension of the data [@Haldane1948-nm]. 
Based on this idea, statistical methods can be used to detect anomalies in investigations. Whether data are in line with the reported probabilistic processes and their theoretically expected outcomes (e.g., random assignment) can indicate deviations from the reported protocol, potentially even data fabrication. 
<!-- ref naar wagenaar 1972? in l21 -->

Statistical methods have proven to be of importance in initiating data fabrication investigations or in assessing scope of potential data fabrication. 
For example, Kranke, Apfel, and Roewer skeptically perceived Fuji's "incredibly nice" data [@10.1213/00000539-200004000-00053] and used statistical methods to contextualize their skepticism.
At the time, a reviewer perceived them to be on a "crusade  against Fujii and his colleagues" [@@10.1111/j.1365-2044.2012.07318.x] and further investigation was absent. 
Only when Carlisle extended the systematic investigation to 168 of Fuji's papers [@10.1111/j.1365-2044.2012.07128.x;@10.1111/anae.13650;@10.1111/anae.13126] did events cumulate into an investigation- and ultimately retraction of 183 peer-reviewed papers [@oransky2015;@10.1016/j.ijoa.2012.10.00]. 
In another example, the Stapel case, statistical evaluation of his oeuvre  occurred after he had already confessed to fabricating data [@Levelt2012].
This resulted in 58 retractions [@oransky2015] and cleared all PhD students of wrongdoing [@Levelt2012].

In order to determine whether generic application of statistical methods to detect data anomalies is responsible, their diagnostic value requires further investigation. 
After all, accusations of data fabrication have grave consequences for the people involved [regretfully, the STAP case brings this to the fore very clearly; @10.1038/520600a]. 
Many of these statistical methods to detect data anomalies are quantifications of initial suspicions by researchers.
Hence, until validated, these should be considered proposed methods.
In some cases, convergent evidence is provided by testing the underlying psychological premises in the general population [@10.1080/08989629508573866;@10.1080/08989620212969] but the question remains whether this premise also applies for data fabricators.
Considering we hardly know how researchers might go about fabricating data, this seems especially problematic. 
Known cases provide relatively few and biased insights into the mind of the data fabricator. 
Relatively extensive descriptions in rare and partial autobiogrophical accounts provide little insight into the actual data fabrication process, except for the setting where it might take place [e.g., late at night when no one is around; @stapel_book].

<!-- Carlisle tested whether Fuji's Randomized Clinical Trials (RCTs) baseline measurements were identically distributed across groups. 
If random assignment occurred, this should be the case. 
As such, the $p$-values for group comparisons would be expected to be uniformly distributed because the null hypothesis of identical distributions across groups is true by definition of the randomized design. 
In the Fuji papers, group comparisons showed excessive consistency, resulting primarily in high $p$-values (e.g., .99, .95) and a high mean $p$-value across the comparisons (mean $p$-value of .5 is expected under uniform distribution).
As an illustration, see Table 1, which depicts 10 hypothetical studies containing 100 participants per condition. 
Set 1 depicts true randomized designs; Set 2 depicts fabricated designs. 
The mean $p$-value for the true randomized design Set 1 is `r round(mean(ex1_1tp), 3)`, whereas the fabricated Set 2 has mean $p$-value `r round(mean(ex2_1tp), 3)`.
 -->

<!-- insert code for the r bits in the next paragraph here -->

We present two studies investigating the diagnostic performance of statistical methods to detect data fabrication.
These studies investigate methods to detect data fabrication in summary statistics (Study 1) or in raw data (Study 2). 
In Study 1, we invited researchers to fabricate summary statistics for a set of four anchoring studies, for which we also had genuine data from the Many Labs 1 initiative [[https://osf.io/pqf9r](https://osf.io/pqf9r); @10.1027/1864-9335/a000178].
In Study 2, we invited researchers to fabricate raw data for a Stroop experiment, for which we also had genuine data from the Many Labs 3 initiative [[https://osf.io/n8xa7/](https://osf.io/n8xa7/); @10.1016/j.jesp.2015.10.012]. 
Before presenting these studies, we expand on the theoretical framework of the investigated statistical methods to detect data fabrication.
<!-- We note that the reviewed methods are not exhaustive of all methods available to test for potential data fabrication in empirical data [see also, @buyse1999;@10.7287/peerj.preprints.2400v1;@10.7287/peerj.preprints.2064v1;@sprite-heathers]. -->

# Theoretical framework

In the current paper, we differentiate between statistical methods to detect potential data fabrication based on reported summary statistics or raw data. 
Below, we expand on the theoretical underpininings of these methods and provide sample code to run these, where appropriate (implemented in the `ddfab` package for R). 
<!-- worth thinking about whether this is ethical if methods prove irresponsible -->
For summary statistics, we review $p$-value analysis, variance analysis, and effect size analysis as potential ways to detect data fabrication. 
$P$-value analyses can be applied whenever a set of nonsignificant $p$-values are reported; variance analysis can be applied whenever a set of variances and accompanying sample sizes are reported for independent, randomly assigned groups; effect size analysis can be used whenever the effect size is reported or can be computed [e.g., an APA reported t- or F-statistic; @10.1525/collabra.71]. 
For raw data, we review digit analyses (i.e., the Newcomb-Benford law and terminal digit analysis) and multivariate associations as potential ways to detect data fabrication. 
The Newcomb-Benford law can be applied when untruncated ratio- or count scale measures are present [@10.1016/j.spa.2005.05.003]; terminal digit analysis can be applied whenever there are sufficient digits [see also @10.1080/08989629508573866]. 
Multivariate associations can be investigated whenever there are two or more variables available and data on that same relation is available from (assumably) genuine data sources.

## Detecting data fabrication in summary statistics

### P-value analysis

The distribution of a single- or a set of $p$-values is uniform if the null hypothesis is true; it is right-skewed if the alternative hypothesis is true [@fisher1925]. 
The distribution of one $p$-value is the result of the population effect size, the precision of the estimate, and the observed effect size, whose properties carry over to a set of independent $p$-values if those $p$-values are independent. 
As such, the $p$-value distribution of a set of independent $p$-values is uniform when the null hypothesis is true, or right-skewed when the alternative hypothesis is true.

When assumptions underlying the computation of a $p$-value are violated, $p$-value distributions can take on a variety of shapes. 
For example, when optional stopping occurs and the null hypothesis is true, $p$-values just below .05 become more frequent [@10.1080/17470218.2014.982664;@10.7717/peerj.1935]. 
However, when optional stopping occurs under the alternative hypothesis or when other researcher degrees of freedom are used, a right-skewed distribution for significant $p$-values can still occur [@10.1037/xge0000086;@10.7717/peerj.1935/fig-1].

When independent $p$-values are not right-skewed or uniformly distributed (as would be theoretically expected), it can indicate potential data fabrication. 
For example, in the Fuji case, supposedly randomly assigned groups were fabricated. 
In truly randomly assigned groups, the measurements are statistically identical (prior to an intervention). 
However, in the Fuji case Carlisle observed many large $p$-values, which ultimately led to the identification of potential data fabrication [@]. 
In Table xxxx we illustrate the difference between expected data under the null distribution (Set 1) and excessively consistent and potentially fabricated data (Set 2). 
More specifically, the expected value of a uniform $p$-value distribution is .5, but the fabricated data from our illustration have a mean $p$-value of `r round(mean(ex2_1tp), 3)`.

```{r table_excess, echo = FALSE}
knitr::kable(df, digits = 3, caption = "Examples of means and standard deviations for a continuous outcome in genuine- and fabricated randomized clinical trials. Set 1 (S1) is randomly generated data under the null hypothesis of random assignment (assumed to be the genuine process), whereas Set 2 (S2) is generated under excessive consistency with equal groups. Each trial condition contains 100 participants. The $p$-values are the result of independent $t$-tests comparing the experimental and control conditions within each respective set.")
```

In order to test whether a distribution of independent $p$-values might be fabricated, we previously proposed using the Fisher method [@fisher1925;@10.1186/s41073-016-0012-9]. 
The Fisher method originally was intended as a meta-analytic tool, which tests whether there is sufficient evidence for an effect (i.e., right-skewed $p$-value distribution). 
This test is computed as
$$
\chi^2_{2k}=-2\sum\limits^k_{i=1}\ln(p_i)
$$
where it tests for more smaller $p$-values than larger $p$-values across the $k$ number of $p$-values. 
Reversing this results in
$$
\chi^2_{2k}=-2\sum\limits^k_{i=1}\ln(1-\frac{p_i-t}{1-t})
$$
where it now tests for more larger $p$-values than smaller $p$-values across the $k$ number of $p$-values that fall above the threshold $t$ (i.e., the Fisher method now tests for left-skew). 
When $t=0$, all $p$-values are selected. 
When $t>0$ the remaining $p$-values are rescaled to fit the original 0-1 range (i.e., dividing by $1-t$). 
This test is similar (but not equivalent) to Carlisle's method testing for excessive homogeneity across baseline measurements in RCTs [@10.1111/anae.13938;@10.1111/j.1365-2044.2012.07128.x;@10.1111/anae.13126].

```{r, echo = FALSE}
threshold <- .05
genuine_p <- ex1_1tp[ex1_1tp > threshold]
fab_p <- ex2_1tp[ex2_1tp > threshold]
genuine_chi <- -2 * sum(log(1 - ((genuine_p - threshold) / (1 - threshold))))
fab_chi <- -2 * sum(log(1 - ((fab_p - threshold) / (1 - threshold))))
genuine_chi_p <- pchisq(q = genuine_chi, df = 2 * length(genuine_p), lower.tail = FALSE)
fab_chi_p <- pchisq(q = fab_chi, df = 2 * length(fab_p), lower.tail = FALSE)
x <- c(.21, -.08, -.37, -.08, .32)
pval <- pnorm(abs(x), 0, 1, lower.tail = FALSE) * 2
pval <- pval[pval > threshold]
chi <- -2 * sum(log(1 - ((pval - threshold) / (1 - threshold))))
p_chi <- pchisq(q = chi, df = 2 * length(pval), lower.tail = FALSE)
```

As an example, we apply the reversed Fisher method to both the genuine- and fabricated results. 
Using the threshold $t=`r threshold`$ to only select the nonsignificant results from Table xxxx, we retain $k=`r length(genuine_p)`$ genuine $p$-values and $k=`r length(fab_p)`$ fabricated $p$-values. 
This results in $\chi^2_{2\times`r length(genuine_p)`}=`r round(genuine_chi, 3)`,p=`r round(genuine_chi_p, 3)`$ for the genuine data from Table xxxx; $\chi^2_{2\times`r length(fab_p)`}=`r round(fab_chi, 3)`,p=`r round(fab_chi_p, 3)`$ for the fabricated data from Table xxxx. 
Another more practical example directly from the Fuji case [@10.1111/j.1365-2044.2012.07128.x], anecdotally illustrates that actual fabricated data can result in significant findings with the reversed Fisher method. 
For example, $p$-values extracted from the original Table 3 [fentanyl dose; @10.1111/j.1365-2044.2012.07128.x] for five independent comparisons show excessively high $p$-values, $\chi^2_{2\times`r length(pval)`}=`r round(chi, 3)`, p=`r round(p_chi, 3)`$.

We note that misspecified one-tailed tests can also result in excessive amounts of large $p$-values.
For correctly specified one-tailed tests, the $p$-value distribution is right-skewed if the alternative hypothesis is true. 
When the alternative hypothesis is true, but the effect is in the opposite direction of the hypothesized effect (e.g., a negative effect when a one-tailed test for a positive effect is conducted), this results in a left-skewed $p$-value distribution.
As such, any data fabrication detected with this method would need to be inspected for misspecified one-tailed hypotheses to preclude false conclusions. 
In the studies we present in this manuscript, misspecification of one-tailed hypothesis testing is not an issue because we prespecified the effect and its direction to the participants.

### Variance analysis

Sample variance- or standard deviation estimates are typically reported to indicate dispersion, but just like the mean there should be sampling error in this estimate proportional to the sample size [i.e., $\sigma/\sqrt{2n}$ under the assumption of normality, p. 351,@yule1922]. 
If an observed random variable $x$ is normally distributed, the variance of $x$ is $\chi^2$-distributed [p. 445; @hogg-tanis]; that is
$$
var(x)\sim\frac{\chi^2_{N_j-1}}{N_j-1}
$$
where $N$ is the sample size of the $j$th group. 
Using the reported summary statistics, we can compute the Mean Squares within ($MS_w$) each condition as
$$
MS_w=\frac{\sum\limits^k_{j=1}(N_j-1)s^2_j}{\sum\limits^k_{j=1}(N_j-1)}
$$
where $s^2_j$ is the variance in the $j$th group. 
The reported variances can be standardized by dividing the observed variances by the $MS_w$; we denote standardized variances with $z^2$.

By standardizing the reported variances observed dispersion across measures can be calculated. 
Observed dispersion of the standardized variances can be operationalized as the standard deviation of the variances [denoted in this paper as $SD_z$, @10.1177/0956797613480366] or as the range of the variances (denoted as $max-min_z$). 
Note that such comparisons are only meaningful if the reported variances originate from a homogeneous population distribution of variances (otherwise, subgroups need to be made).

In order to compute the expected dispersion, we use the distribution of the standardized variances to compute how extreme the observed dispersion of the variances is. 
Assuming normality, the distribution of the standardized variances follows a $\chi^2$-distribution in the form of
$$
z^2_j\sim\left(\frac{\chi^2_{N_j-1}}{N_j-1}\right)/MS_w
$$
which is the result of the distribution of $var(x)$ as denoted earlier, divided by the $MS_w$. 
From this distribution, fictitious variances can be generated for each group under investigation. These are subsequently used to compute one of the measures for dispersion proposed in the previous paragraph (e.g., standard deviation of the variances). 
Repeating this across $i$ iterations provides an estimate of the density function for the expected dispersion of the variances. By comparing the observed dispersion of the variances with the expected variances, we can estimate how extreme our observations are. 
For our purposes, too little dispersion in the variances indicates potential fabrication in the reported data [@10.1177/0956797613480366].

```{r, echo = FALSE}
sd1 <- c(ex1_1esd, ex1_1csd)
sd2 <- c(x2_1esd, x2_2csd)

iterate <- 10000
res1 <- std_var(n = rep(100, length(sd1)), sds = sd1, iter = iterate)
res2 <- std_var(n = rep(100, length(sd2)), sds = sd2, iter = iterate)
```

As an example, we apply the variance analysis to the illustration from Table xxxx and the Smeesters case. 
For the reported standard deviations in Table xxxx, we apply the variance analysis across those in the genuine- and fabricated sets separately. 
For the genuine data (Set 1), we find that the reported mean standard deviation is `r mean(sd1)` with a standard deviation of `r sd(sd1)`; for the fabricated data (Set 2), we find that the reported mean standard deviation is `r mean(sd2)` with a standard deviation of `r sd(sd2)`. 
These summary statistics of the summary statistics already indicate there is a difference between the genuine- and fabricated data.
Variance analysis helps us quantify this: Set 1 has no excessive consistency (`r round(res1, 3)`), whereas Set 2 does show excessive consistency (`r round(res2, 3)`). 
In words, out of `r iterate` theoretically expected samples, `r res1 * iterate` showed more consistency for Set 1, whereas only `r res2 * iterate` showed more consistency for Set 2. 
As a non-fictional example, three independent conditions from the same study ($n_k=15$) were reported to have standard deviations 25.09, 24.58, and 25.65 in the Smeesters case. 
The standard deviation of these standard deviations is `r round(sd(c(25.09, 24.58, 25.65)), 2)` (i.e., $SD_z$). 
Such consistency (or even more) would only be observed in `r round(std_var(n = rep(15, 3), sds = c(25.09, 24.58, 25.65), iter = iterate) * 100, 2)`% of 100,000 simulated replications [@10.1177/0956797613480366].

# Study 1 - detecting fabricated summary statistics

We tested the performance of statistical methods to detect data fabrication in summary statistics with genuine- and fabricated summary statistics from four anchoring studies [@10.1126/science.185.4157.1124;@10.1037/e722982011-058]. 
The anchoring effect is a well-known psychological heuristic that uses the information in the question as the starting point for the answer, which is then adjusted to yield a final estimate of a quantity. 
For example 'Do you think the percentage of African countries in the UN is above or below  [10\% or 65\%]?  What do you think is the percentage of African countries in the UN?'. 
These questions yield mean responses of 25\% and 45\%, respectively [@10.1126/science.185.4157.1124], despite essentially posing the same factual question. 
A considerable amount of (assumably) genuine data sets on the anchoring  heuristic are freely available [[https://osf.io/pqf9r](https://osf.io/pqf9r); @10.1027/1864-9335/a000178] and we collected fabricated data sets within this study. 
This study was approved by the Tilburg Ethical Review Board (EC-2015.50).

## Methods

We collected summary statistics for four anchoring studies: (i) distance from San Francisco to New York, (ii) population of Chicago, (iii) height of the Mount Everest, and (iv) the number of babies born per day in the United States [@10.1037/e722982011-058]. 
Each of the four studies provided us with summary statistics for a 2 (low/high anchoring) $\times$ 2 (male/female) factorial design. 
Throughout our study, the unit of analysis is a set of summary statistics (i.e., means, standard deviations, and test results) for the four anchoring studies from one respondent. 
For current purposes, a respondent is defined as researcher/lab where the four anchoring studies' summary statistics originate from. 
All materials, data, and analyses scripts are freely available on the OSF ([https://osf.io/b24pq](https://osf.io/b24pq)) and a preregistration is available at [https://osf.io/tshx8/](https://osf.io/tshx8/). 
Throughout this report, we will indicate which facets were not preregistered or deviate from the preregistration by denoting "(not preregistered)" or "(deviation from preregistration)", respectively.

### Data collection

We downloaded thirty-six genuine data sets from the publicly available Many Labs (ML) project [[https://osf.io/pqf9r](https://osf.io/pqf9r); @10.1027/1864-9335/a000178]. 
The ML project replicated several effects across thirty-six locations, including the anchoring effect in the four studies mentioned previously. 
Considering the size of the ML project, the transparency of research results, and minimal individual gain for fabricating data, we assumed these data to be genuine. 
For each of the thirty-six locations we computed three summary statistics (i.e., sample sizes, means, and standard deviations) for each of the four conditions in the four anchoring studies (i.e., $3\times4\times4$) for each of the thirty-six locations. 
We computed these summary statistics from the raw ML data, which were cleaned using the original analysis scripts from the ML project.

# References